\documentclass{article}

\usepackage{amsmath}
\usepackage{unicode-math}

\DeclareMathOperator{\Bij}{Bij}
\renewcommand{\k}{\mathbb{k}}

\title{Towards manifolds}

\author{Patrick Massot}

\begin{document}
\maketitle

\section{Algebraic preliminaries}
\label{sec:algebraic_preliminaries}

A left action of a group $G$ on a set $X$ is a group morphism $ρ$ from $G$
to the group of permutation $\Bij(X)$. The orbit $Gx$ of a point $x ∈ X$
under $G$ is $\{ ρ(g)(x); g ∈ G\}$. An action is transitive if 
$Gx = X$ for some (or equivalently all) $x$ in $X$. An action $ρ$ is
called free when it is injective. A principal $G$-space is a set
equipped with a transitive free action of $G$.

An affine space over a field $\k$ is principal $E$-space $X$ for some
$k$-vector space $E$. We also say that $X$ is directed by $E$.
Notations: the action of $v ∈ E$ on a point 
$x ∈ X$ is denoted by $x + v$. Given $x$ and $y$, the unique $v$ in $E$
such that $y = x + v$ is denoted by $y - x$.

One could also define affine spaces without going through the general
definition of group actions (spelling out everything in this special
case).

\section{Topological preliminaries}
\label{sec:topological_preliminaries}

A topological vector space $E$ over a topological field $\k$ is a
$\k$-vector space equipped with a topology such that addition is
continuous from $E × E$ to $E$ and multiplication is continuous from 
$\k × E$ to $E$.

For every affine space directed by a topological vector space, there is
a unique topology such that each $v ↦ x + v$ is a homeomorphism.

Let $\k$ be a field. An absolute value on $\k$ is a map $|·|$ from 
$\k$ to $ℝ$ such that $|x| = 0 ⇔ x = 0$, $|xy|= |x|·|y|$, 
$|x + y| ≤ |x| + |y|$.
A norm on a $\k$-vector space $E$ with absolute value is a map $‖·‖$
from $E$ to $ℝ$ such $‖x‖ = 0 ⇔ x = 0$, $‖x + y‖ ≤ ‖x‖ + ‖y‖$, 
$‖λ x‖ =|λ| · ‖x‖$.

Any two norms $‖·‖₁$ and $‖·‖₂$ on a finite dimensional real
vector space are equivalent: there exists a positive real number $C$
such that: $‖·‖₁/C ≤ ‖·‖₂ ≤ C‖·‖₁$ (this is not quite trivial, requires
compactness…). Although this result could be avoided, it would make
everything clumsy, carrying out irrelevant choices of norms everywhere.

To each norm $‖·‖$ one can associate a distance $d(x, y) = ‖x - y‖$, and
then associated topology. In the case of finite dimensional vector
spaces, the resulting topology is independent of the chosen norm (by
equivalence of norms), and all linear maps are continuous.

Norm on continuous linear maps spaces $L(E, F)$: 
$‖φ‖ = \max \{‖φx‖ ; ‖x‖ = 1\}$. This makes $L(E, F)$ into a normed
algebra (normed vector space and $‖φ ∘ ψ‖ ≤ ‖φ‖·‖ψ‖$). 

\section{Differential calculus}
\label{sec:differential_calculus}

Let $X$ and $Y$ be two affine space directed by normed vector spaces
$E$ and $F$. Let $U$ be an open subset of $X$. A map $f : U → Y$ is
differentiable at $x ∈ U$ if there exists a continuous linear map 
$L : E → F$ such that, for sufficiently small $v$,
$f(x + v) = f(x) + Lv + R(v)$ where $‖R(v)‖ = ‖v‖ε(v)$ for some function 
$ε : E → ℝ$ which goes to zero when $v$ goes to $0$ (let's pretend it
would be reasonable not to define $o(v)$ notation). In this case, $L$ is
unique, and denote by $Df(x)$. The map $f$ is differentiable on $X$ if
it is differentiable at every $x$ in $X$. By induction, define
$(k+1)$-times differentiable maps by $Df$ is $k$-times differentiable.
It is $C¹$ if $x ↦ Df(x)$ is continuous. By induction, a map is
$C^{k+1}$ if $Df$ is $C^k$. It is smooth, or $C^∞$ if it is $C^k$ for
all $k$. 

The chain rule. Let $X$, $Y$ and $Z$ be affine spaces. Let $U$ and $V$
be open subsets in $X$ and $Y$ respectively. If $f : U → V$ is
differentiable at $x$ and $g : V → Z$ is differentiable at $f(x)$ then
$g ∘ f$ is differentiable at $x$, and $D(g ∘ f)(x) = Dg(f(x)) ∘ Df(x)$.

A composition of two $C^k$ maps is $C^k$ (but let's not state the
formula for the higher order differential).

Almost any serious proof in differential topology will rely on the
inverse function theorem, but not the early definitions so let's omit it
at first.
\end{document}
